---
title: "p8105_hw2_jl6048"
output: github_document
author: Jinghan Liu
---

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1a

#Read and clean the Mr. Trash Wheel sheet
```{r message=FALSE}
Trash_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel" ,range ="A2:N535") %>%
  janitor::clean_names() %>%
  drop_na(dumpster,month) %>%
  mutate(sports_balls = round(sports_balls)) 
Trash_df
```


## Problem 1b

#Read and clean precipitation data for 2018 and 2019
```{r message=FALSE}
data2019_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation" ,range ="A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2019)
data2019_df

data2018_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation" ,range ="A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018) 
data2018_df
```

#combine precipitation datasets
```{r message=FALSE}
data_df =
  bind_rows(data2019_df, data2018_df)%>%
  mutate(month = month.name[month])
data_df
```

#description

The number of obervations in Trash wheel dateset is `r nrow(Trash_df)` obsevations and the varialbles are `r ncol(Trash_df)` variables. 
The key variables for trash measurements are `r Trash_df %>% select(weight_tons)%>% names()`

The number of obervations in combined precipitation dateset is `r nrow(data_df)` obsevations and the varialbles are `r ncol(data_df)` variables. 
The total precipitation in 2018 was`r sum(pull(data2018_df,total))` inches
The median number of sports balls in a dumpster in 2019 was`r Trash_df %>% filter(year == 2019) %>% pull(sports_balls)  %>% median()`



## Problem 2

#pols dataset read and tidy
```{r message=FALSE}
pols_df = 
  read_csv("data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(month = month.abb[month])
pols_df


pols_tidy =
  pols_df %>%
  pivot_longer(
    c(prez_gop, prez_dem),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "dbi") %>%
    select(-day) %>%
  filter(dbi != 0) %>%
    select(-dbi)
pols_tidy

```
#snp dataset read and tidy
```{r message=FALSE}
snp_df = 
  read_csv("data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month","day","year"), convert = TRUE)%>%
   mutate(
     year = ifelse(year<= 15, year + 2000, year + 1900),
     month = month.abb[month]) %>%
  select(year, month, close)
snp_df
```
#unemployment dataset read and tidy
```{r message=FALSE}
une_df = 
  read_csv("data/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment")%>%
  janitor::clean_names() 
une_df
 
```
#merge three datasets
```{r message=FALSE}
pols_snp_df =
  full_join(pols_tidy, snp_df)
all_df =
  full_join(pols_snp_df, une_df)
all_df
```

#description

The pols-months dataset have 822 observations and 9 variables. This data contains 1947-2015 years national politicians and in the column of President, "dem" means democratic and "gop" means republican.

The snp dataset have 787 observations and 3 variables. This data contains the closing values of the standard & poor's stock index in 1950-2015. The largest closing value was`r max(pull(snp_df,close))` and the smallest closing value was `r min(pull(snp_df,close))`

The unemployment dataset have 816 observations and 3 variables. This dataset contains unemplyment percentage from `r min(pull(une_df, year))` to `r max(pull(une_df, year))`.



## Problem 3

#remove duplicate rows
```{r}
pbn_df = 
  read_csv("Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate( gender = str_to_lower(gender), 
          childs_first_name = str_to_lower(childs_first_name),
          ethnicity = str_to_lower(ethnicity),  
          ethnicity = recode(ethnicity, "white non hisp" = "white non hispanic", "asian and paci" = "asian and pacific islander", "black non hisp" = "black non hispanic"))%>%
  distinct()
pbn_df
```

#Olivia's table
```{r message=FALSE}
olivia_df = 
  filter(pbn_df,
    gender == "female",
    childs_first_name == "olivia") %>%
  select(year_of_birth, ethnicity, rank) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>%
  knitr::kable()
olivia_df
```

#males table
```{r message=FALSE}
male_df = 
  filter(pbn_df,
    gender == "male",
    rank == 1) %>%
  select(year_of_birth, ethnicity, childs_first_name) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>%
  knitr::kable()
male_df
```

# scatter plot 
```{r message=FALSE}
male_white_2016 =
  filter(pbn_df,
         gender == "male",
         year_of_birth == "2016",
         ethnicity == "white non hispanic")%>%
  ggplot(aes(x = rank, y = count)) +
  geom_point() + 
  geom_text(aes(label = childs_first_name)) +
  labs(title = "white non-hispanic male children name rank in 2016",
       x = "Rank of name",
       y = "number of name")
male_white_2016
```

